Extending Height Above Nearest Drainage to Model Multiple Fluvial Sources in Flood Inundation Mapping Applications for the U.S. National Water Model

### 2022-09-15 manuscripts/owp_fim4_2022/sections/materials_and_methods.tex
* latest updates * possible resubmission candidate

For a more intimate look at the drainage order reduction procedure and its effects, GMS, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.

@@

\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into LPs (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.

@@

Utilizing the stage-discharge relationships in the SRCs, each forecast for each catchment identifier is assigned a stage value. 
The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
%
\begin{linenomath*}
\begin{equation}
\label{eq:hand_fim_depth}
    D_{ij} = S_i - H_{ij}
\end{equation}
\end{linenomath*}
%
\begin{linenomath*}
\begin{equation}
\label{eq:hand_fim}
    F_{ij} = D_{ij} > 0
\end{equation}
\end{linenomath*}
%
For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
Consolidating the depths using a maximum function was decided upon based on intuition which we believe to best represent the depth of water in an area with multiple contributing fluvial inundation sources.
Other aggregation methods could lead to different results but were not investigated here.
Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i,
%
\begin{linenomath*}
\begin{equation}
\label{eq:comp_fim_depths}
    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
\end{equation}
\end{linenomath*}
%
. Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps,
%
\begin{linenomath*}
\begin{equation}
\label{eq:comp_fim}
    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
\end{equation}
\end{linenomath*}
. For the MS and GMS methods, the contributing areas are defined differently.
For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to two for that case.
For GMS, all FIMs from all the LPs in a given area are mosaiced together then K is set to this number of LPs.
Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}
\label{ssec:evaluation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Possible benchmark FIM candidates for evaluation purposes include high-water marks, remote sensing observation, crowd-sourced information, and modeled extents.
Unfortunately, all of these sources or using a combination of them are subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to isolating out hydrological factors.
A suitable dataset needed to be selected to best balance advantages and limitations with future studies interested in expanding to additional datasets as able to.

Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D derived models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, more sophistication when compared to HAND based FIMs, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
49 HUC8s spanning about 185 thousand $km^2$ were available at the time (now more) across nine states and shown in Figure \ref{fig:all_ble_maps}.
The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM and thus corresponding discharges and mapping extents for evaluation.
We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.

By using the same HEC-RAS derived discharges and FIM extents, we are able to separate out errors introduced by hydrology, atmospheric forcings, hydraulic routing, and others that we would have potentially affected our conclusions if we used NWM forecasted discharges.
Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with Horton-Strahler orders of one due to the lack of representation of these additional headwater streams in the NWM network.
While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
%
\begin{figure}[H]
\centering
\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
\caption{
Shows 185 thousand $km^2$ of modeled areas for the Base Level Engineering (BLE) domain of 49 HUC8s across nine states at 0.2\% recurrence magnitude for flow rates.
BLE maps are produced for two recurrence flows, 1\% (100 yr) and 0.2\% (500 yr), using 1D HEC-RAS models.
The maps are used as benchmarks for validation purposes of OWP FIM.
}
\label{fig:all_ble_maps}
\end{figure}
%
\begin{figure}[H]
\centering
\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
\caption{
Illustrates Base Level Engineering (BLE) cross sections and stream lines at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowline.
The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
This creates additional inundation areas in the validation data that are not modeled with our HAND based FIMs.
}
\label{fig:ble_evaluation_method}
\end{figure}
%

>>>>

For a more intimate look at the drainage order reduction procedure GMS, and its effects, we allude to Figure \ref{fig:gms_methods_2} which references the same area (in HUC8 12090301) and set of river junctions as in Figure \ref{fig:catchment_boundaries_issue}.

@@

\caption{Illustrates the NWM Full Resolution V2.1 stream network discretized into level paths (LP), symbolized by unique colors, as well as the values of the arbolate sums to the nearest whole km distance.

@@

Utilizing the stage-discharge relationships in the SRCs, each forecast for each catchment identifier is assigned a stage value. 
The catchments grid encoded with the reach identifiers are used to map the stages by thresholding to the forecast stage.
We use the basic logic already established in previous works to conduct this \cite{nobre2016hand,liu2016cybergis,maidment2017conceptual}.
Mathematically, the HAND values, $H_{ij}$, can be indexed by the reach identifiers, i, and pixel indices, j.
For each forecast stage, $S_i$, one can express the formula for $D_{ij}$, a continuous variable denoting water depth at a given pixel with reach and pixel identifiers i and j respectively in Equation \ref{eq:hand_fim_depth}.
For each forecast stage, $S_i$, one can express the formula for $F_{ij}$, a binary variable denoting inundation condition in Equation \ref{eq:hand_fim} in terms of $D_{ij}$ by simply thresholding at zero depths.
%
\begin{linenomath*}
\begin{equation}
\label{eq:hand_fim_depth}
    D_{ij} = S_i - H_{ij}
\end{equation}
\end{linenomath*}
%
\begin{linenomath*}
\begin{equation}
\label{eq:hand_fim}
    F_{ij} = D_{ij} > 0
\end{equation}
\end{linenomath*}
%
For the cases of MS and GMS, the inundation maps produced for the respective processing units at lower maximum stream orders must be mosaiced together to form a seamless forecast in the form of a single raster file.
For mosaicing the depths, we select the maximum inundation depth from the all the contributing areas K index by its lower case character, k.
Consolidating the depths using a maximum function was decided upon based on intuition which we believe to best represent the depth of water in an area with multiple contributing fluvial inundation sources.
Other aggregation methods could lead to different results but were not investigated here.
Equation \ref{eq:comp_fim_depths} illustrates how the maximum depth from all the contributing areas, k, to each pixel j in catchment i,
%
\begin{linenomath*}
\begin{equation}
\label{eq:comp_fim_depths}
    D_{ij} = \max_{k=[1,...,K]} D_{ijk}
\end{equation}
\end{linenomath*}
%
. Equation \ref{eq:comp_fim} illustrates the same process but for mosaicing the binary inundation maps,
%
\begin{linenomath*}
\begin{equation}
\label{eq:comp_fim}
    F_{ij} = \max_{k=[1,...,K]} F_{ijk}
\end{equation}
\end{linenomath*}
. For the MS and GMS methods, the contributing areas are defined differently.
For MS, the FIM from MS HAND and FR HAND are mosaiced together to form a singular inundation map thus K is set to two for that case.
For GMS, all FIMs from all the LPs in a given area are mosaiced together then K is set to this number of LPs.
Figures \ref{fig:gms_methods}a and \ref{fig:gms_methods}b, illustrate how inundation maps are created for lower stream order processing units then mosaiced together.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evaluation}
\label{ssec:evaluation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Possible benchmark FIM candidates for evaluation purposes include high water marks, remote sensing observation, crowd-sourced information, and modeled extents.
These sources are all subject to limitations for evaluating a continental scale model like OWP FIM such as but not limited to a lack of spatial coverage, signal interference, lack of streamflow data, inaccurate streamflow data, physics-based assumptions, and errors in input data.
While in-situ observations such as high water marks offer the highest accuracy, they are often limited in spatial extent and can lack the associated streamflow data necessary to make FIMs to compare to as to isolate out other hydrological factors.

Evaluation of our relative elevation CFIM method is conducted by comparison to the HEC-RAS 1D models produced within FEMA region 6 \cite{fema2021base,fema2021estimated,us2022hydrologic}.
This dataset was selected due to its large spatial coverage, availability of cross-sections with streamflow information, higher level of sophistication when compared to HAND, engineering scale detail, and a storied use in the literature as an evaluation dataset \cite{cook2009effect,rajib2016large,zheng2018geoflood,afshari2018comparison,wing2017validation,criss2022stage,follum2017autorapid}.
We selected 49 available HUC8s, shown in Figure \ref{fig:all_ble_maps}, which span about 185 thousand $km^2$ across nine states.
The maps of the 1\% recurrence flow (1 in 100 year) and the 0.2\% recurrence flow (1 in 500 year) are furnished by InFRM as well as the corresponding discharges and mapping extents for evaluation.
We did exclude NWM V2.1 Reservoirs from evaluation because these are not properly accounted for in the inundation sourced from OWP FIM.

By using the same HEC-RAS derived discharges and FIM extents for creating maps with OWP FIM, we are able to separate out errors introduced by NWM inputs and processes including land surface interactions, groundwater fluxes, atmospheric forcings, hydraulic routing, and others that would have potentially affected our conclusions if we had used NWM forecasted discharges.
Figure \ref{fig:ble_evaluation_method} illustrates both NWM V2.1 and BLE stream lines as well as the BLE cross-sections that have recurrence discharges associated with them.
We elected to spatially intersect the HEC-RAS cross sections with the NWM stream network assigning the 1\% and 0.2\% flow rates to each NWM reach. 
To handle multiple intersections, we opted to use a filter to select the median discharge value attributed to each NWM reach.
This partially handles the influence of neighboring cross sections that could cause flow discontinuities and mass conservation issues.
Additionally, the stream network of the InFRM furnished models are of higher stream densities and bifurcation ratios, as evident in Figure \ref{fig:ble_evaluation_method}, leading to a significant amount of false negatives (FN) (under-prediction) along headwater streams with unit Horton-Strahler order due to the lack of representation of these additional headwater streams in the NWM network.
While the limitations are noted, this method does best to detangle the influence of exogenous variables that we do not wish to study in this comparison.
%
\begin{figure}[H]
\centering
\includegraphics[scale=1.0]{figures/all_ble_maps.jpg}
\caption{
Shows 185 thousand $km^2$ of modeled areas for the Base Level Engineering (BLE) domain of 49 HUC8s across nine states at 0.2\% recurrence magnitude for flow rates.
BLE maps are produced for two recurrence flows, 1\% (100 yr) and 0.2\% (500 yr), using 1D HEC-RAS models.
The maps are used as benchmarks for validation purposes of OWP FIM.
}
\label{fig:all_ble_maps}
\end{figure}
%
\begin{figure}[H]
\centering
\includegraphics[scale=1.0]{figures/ble_evaluation_method.jpg}
\caption{
Illustrates Base Level Engineering (BLE) cross sections and stream lines at the HUC8 12100203 near the confluences of West Fork Plum Creek and Clear Fork Plum Creek with Plum Creek.
BLE cross sections are intersected with NWM reaches and the median recurrence discharge for 1\% and 0.2\% levels are selected per NWM V2.1 Full Resolution (FR) stream lines.
Additionally, we illustrate the NWM V2.1 catchments to provide a sense of how many cross-sections may intersect a given NWM flowline.
The BLE stream network is also shown which is denser than the NWM V2.1 stream lines meaning there are several lower order streams represented in the BLE stream network that are not in the NWM V2.1 stream lines.
This creates additional inundation areas in the validation data that are not modeled with our HAND based FIMs.
}
\label{fig:ble_evaluation_method}
\end{figure}
%

### 2022-09-15 manuscripts/owp_fim4_2022/sections/openresearch.tex
* latest updates * possible resubmission candidate

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\openresearch
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\openresearch
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
National Water Model (NWM) data used in this study includes the hydrofabric related datasets \cite{nwm2022hydrofabric} including catchments, streamlines, and reservoirs \cite{nwm2022hydrofabric}.
These are furnished by the National Oceanic and Atmospheric Administration (NOAA) Office of Water Prediction (OWP) via an Earth Science Information Partners (ESIP) Amazon Web Services (AWS) S3 Bucket \cite{esipData2022}.
OWP Flood Inundation Mapping (FIM) capabilities rely extensively on the use of the National Hydrography Plus High Resolution (NHDPlusHR) datasets including BurnLineEvents \cite{nhdplus2022vectors}, value-added attributes (VAA) \cite{nhdplus2022vectors}, water boundaries (WBD) or hydrologic unit code (HUC) geometries \cite{nhdplus2022wbd}, and digital elevation maps (DEM) \cite{nhdplus2022dems}.
Some additional datasets for processing include the National Levee Database (NLD) furnished by the United States Army Core of Engineers (USACE) \cite{engineers2016national}, Land-sea border from the Great Lakes Hydrography Dataset (GLHD) furnished by the Great Lakes Aquatic Habitat Framework (GLAHF) \cite{GreatLakesHydrographyDataset}, and a Land-sea border provided by OpenStreetMap (OSM) \cite{osm2021landsea}.
Evaluation data was furnished by Interagency Flood Risk Management (InFRM) consortium including cross-sections and flood depths \cite{fema2021base,fema2021estimated}.
Additionally, some FIM hydrofabric data including HAND grids, catchments, streamlines, synthetic rating curves, and cross-walk tables are available on the ESIP bucket \cite{esipData2022}.

### 2022-09-15 manuscripts/owp_fim4_2022/sections/results.tex
* latest updates * possible resubmission candidate

The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), given Manning's n value (0.06, 0.12), and given recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
This sub-figure shows the CSI distributions across all 49 HUC8s when Manning's n is set to 0.06.
Each independent, whole violin represents the distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes together as they tend to observe similar trends across the three models.
The slope of each trend line is quantified in the figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).

% Interpretation of metrics
Both Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} contain a fair amount of information that is central to the objectives of this study.
As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
This increase is primarily driven by an increase in POD thus generally increasing the probability of correctly detecting inundation.
Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
So overall and in other words, the entire distribution of HUC8s improves across the three methods.
Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
We move more details on the relationship between stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).

@@

The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value used in the BLE model, 0.11 \cite{fema2021base,fema2021estimated}.
Most of this improvement is driven by significant increases in POD but, unfortunately, leading to a significant amount of over-prediction as observed by the increase in FAR.

@@

\includegraphics[scale=0.9]{figures/violin_plots.jpg}

@@

Trend lines for each metric - Mannings combination are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}

@@

@@

For the MS resolution an additional, 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.

>>>>

The distribution of these flood extent metrics can be examined in Figure \ref{fig:violin_plot} as violin plots.
Each half of a violin plot represents the kernel density estimation (KDE) for a given model (FR, MS, GMS), Manning's n value (0.06, 0.12), recurrence interval (1\%, 0.2\%), and performance metric (CSI, POD, FAR).
For example, let's examine the violin plot for the row marked CSI and column for Manning's n = 0.06.
This sub-figure shows the CSI distributions across all 49 HUC8s when Manning's n is set to 0.06.
Each independent, whole violin represents the HUC8 metric value distribution of FR, MS, or GMS while each half of the violin represents the distribution of that data divided up by magnitude (blue for 100 yr and orange for 500 yr).
The horizontal dashed and dotted lines represent the 25th, 50th, and 75th percentiles from bottom to top, respectively.
Additionally, we show trend lines symbolized in green that for each metric and Manning's n combination denotes the best fit line for the three methods (FR, MS, and GMS).
To avoid having two trend lines per sub-figure, we elected to aggregate the two magnitudes together as they tend to observe similar trends across the three models.
The slope of each trend line is quantified in the figure by its $\beta_1$ value and the p-value for that statistic which tests the significance of that trend (deviation from a zero sloped trend line).

% Interpretation of metrics
Both Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics} contain a fair amount of information that is central to the objectives of this study.
As previously stated in Section \ref{ssec:evaluation}, we consider CSI as a general proxy for the skill of the inundation extents with POD denoting skill on inundated areas and FAR indicating skill on non-inundated areas.
Again, the main objective of the study is to introduce how computing HAND with disaggregated stream networks to those with unit stream order can enhance the fidelity of FIMs by capturing fluvial inundation from multiple sources as opposed to that of just the nearest drainage line.
As can be seen in Figure \ref{fig:violin_plot} and Table \ref{tab:aggregate_metrics}, CSI generally increases from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
This increase is primarily driven by an increase in POD thus generally increasing the probability of correctly detecting inundation.
Also, we note that FAR is somewhat, albeit marginally, decreased from FR to MS and MS to GMS for both sets of Manning's n values and flood magnitudes.
The increases in CSI and POD as well as the decreases in FAR with respect to the methods, FR, MS, and GMS, are not only observed among the trend lines but also in the 25th, 50th, and 75th percentiles (Figure \ref{fig:violin_plot}).
So overall and in other words, the broader distribution of HUC8s improves across the three methods.
Due to the means by which FIM is produced utilizing FR, MS, and GMS, we can say that the more we derive HAND on networks of unit stream order and mosaic the resulting FIMs, the better those FIMs perform.
We move more details on the relationship between stream order and FIM skill to the Discussion section (Section \ref{sec:discussion}).

@@

The higher Manning's n value enhances performance for both recurrence intervals across all models which seems to better agree with the value of 0.11 used in the BLE model \cite{fema2021base,fema2021estimated}.
Most of this improvement is driven by significant increases in POD, but unfortunately, it also leads to a significant amount of over-prediction as observed by the increase in FAR.

@@

\includegraphics[scale=0.8]{figures/violin_plots.jpg}

@@

Trend lines for each combination of metric and Manning's n are shown (sample size = 294) along with associated slope and p-value of slope testing one-tailed significance.}

@@

%\multirow{2}{*}{Metric} & \multirow{2}{*}{Manning's n} & \multicolumn{2}{|c|}{FR} & \multicolumn{2}{|c|}{\shortstack{MS\\(\% Change)}} & \multicolumn{2}{|c|}{\shortstack{GMS\\(\% Change)}} \\

@@

For the MS resolution, an additional 0.05 CPU years is required on top of this bringing the total to about 0.18 CPU years to produce 2,188 HUC8s that span additional areas not covered in previous HAND versions including Hawaii and Puerto Rico.
